### 2023-04-24 Notes/main.tex
Added optimal a_2 b_2 theory to notes.

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\logistic}{{\text{logistic}}}

@@

  \item However, one of the key steps is that \textit{all} of the response
    models are free of $Y$. In a previous iteration of Simulation 4, we had
    adopted the framework of \cite{robins1997non} where we first to observe the
    first variable and see if we observe the second variable. In this case, the
    second step can depend on the result of the first step and this is what we
    did. However, this makes it the case that $\pi_11$ is a function of $X_i$
    and $Y_1$ and $Y_2$. In this case $\hat \theta_{eff}$ is not unbiased if the
    reponse model is misspecified (or even just estimated).

  \item If we modify Simulation 4, such that the second step of observed the
    second varible is proportional to $\logistic(y_k)$ then we get the same
    result as before:

@@

\section*{Proposal: Full Nonmonotone Estimator}

\begin{itemize}

  \item When constructing
    estimates of the response model, we estimate $\pi_{11}$ using
    $\pi_{A_2 = 1|A_1 = 1}$ and we never include information about $\pi_{A_1 =
    1|A_2 = 1}$. I would like to create a full estimator where we are able to
    use this information.

  \item The current estimator is the following (See Slide 24 of Non-monotone
    Presentation):

    \begin{align*}
      \hat \theta_{eff} &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \\
      &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)} (b_2(X_i, Y_{1i}) - 
      E[g_i \mid X_i]) \\
      &+ n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)} (a_2(X_i, Y_{2i}) -
      E[g_i \mid X_i]) \\
      &+ n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_1, Y_{1i})}(g_i 
      - b_2(X_i, Y_{1i}) - a_2(X_i, Y_{2i}) + E[g_i \mid X_i])
    \end{align*}

  \item The problem with this is that we assume $\pi_{11}(X, Y_{1}) =
    \pi_{2|1} (X, Y_{1}) \pi_{1+}(X)$ when we could have $\pi_{11}(X, Y_2) =
    \pi_{1|2} (X, Y_2) \pi_{2+}(X)$. In other words, the previous result
    implicitly assumes that $\pi_{11}$ is a function of $X$ and $Y_1$ when it
    could be a function of $X$ and $Y_2$ as well.

  \item I think that we could use the
    following (small modification) instead,

    \[\pi_{11}(X, Y_1, Y_2) = \alpha \pi_{1|2} \pi_{2+} + (1 - \alpha) \pi_{2|1}
    \pi_{1+}\]

    for some $\alpha \in [0, 1]$.

  \item The reasoning behind this is two-fold. First, it allows us to think of
    the nonmonotone missingness case as a linear combination of two monotone
    cases, which I think is useful. Second, we can make explicit our choice of
    $\alpha$, which in the existing estimator is simply $\alpha = 0$.

  \item The problem with adding another parameter $\alpha$ is that it makes the
    overall model unidentifiable. We cannot estimate $\alpha$ and the
    conditional and marginal distributions that we have without additional
    assumptions. So for now, I think that we should just assume that $\alpha$ is
    known and then we apply this method to a data integration problem we can
    review what $\alpha$ makes sense or figure out an additional assumption that
    can help us estimate $\alpha$ (for example if there is a variable correlated
    with $1|2$ versus $2|1$).

  \item This is not just a pure Bayes' rule.
    Formally, let $L = (X, Y_1, Y_2)$. By Bayes' rule we have

    \begin{align*}
      \Pr(R_1 &= 1 \mid R_2 = 1, L) \Pr(R_2 = 1 \mid L) \\
      &= \Pr(R_1 = 1, R_2 = 1 \mid L) \\
      & = \Pr(R_2 = 1 \mid R_1 = 1, L) \Pr(R_1 = 1 \mid L).
    \end{align*}

    Yet to estimate this model, we need to assume something like the following:
    \begin{align*}
      \Pr(R_1 &= 1 \mid R_2 = 1, X, Y_2) \Pr(R_2 = 1 \mid X, Y_2)\\
      &= \Pr(R_1 = 1, R_2 = 1 \mid L) \\
      &= \Pr(R_2 = 1 \mid R_1 = 1, X, Y_1) \Pr(R_1 = 1 \mid X, Y_1)
    \end{align*}

    in which case the two sides are clearly different.

    \item Unfortunately, early simulation results are not promising because we
      cannot distinguish points that should have $\pi_{11}$ conditional on $Y_1$
      and points that should be conditional on $Y_2$.

      \input{Tables/alphasim.tex}

    \item Dr. Kim, please let me know what you think about this idea and if it
      is worth pursuing more.

@@

>>>>

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}

@@

  \item However, one of the key steps is that \textit{all} of the response
    models are free of $Y$. In a previous iteration of Simulation 4, we had
    adopted the framework of \cite{robins1997non} where we first to observe the
    first variable and see if we observe the second variable. In this case, the
    second step can depend on the result of the first step and this is what we
    did. However, this makes it the case that $\pi_11$ is a function of $X_i$
    and $Y_1$ and $Y_2$. In this case $\hat \theta_{eff}$ is not unbiased if the
    response model is misspecified (or even just estimated).

  \item If we modify Simulation 4, such that the second step of observed the
    second variable is proportional to $\logistic(y_k)$ then we get the same
    result as before:

@@

\section*{Minimizing the Variance}

\begin{itemize}
  \item The goal of this section is to find optimal values of $b_2(X, Y_1)$ and
    $a_2(X, Y_2)$ such that the variance of $\hat \theta_{eff}$ is minimized.

  \item Recall:

    \begin{align*}
      \hat \theta_{eff} - \hat \theta_n 
      &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \left(1 - \frac{R_{1i}}{\pi_{1+}}
      - \frac{R_{2i}}{\pi_{2+}} + \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \\
      &+ n^{-1} \sum_{i = 1}^n b_2(X_i, Y_{1i}) \left(\frac{R_{1i}}{\pi_{1+}} -
      \frac{R_{1i}R_{2i}}{\pi_{11}}\right)\\
      &+ n^{-1} \sum_{i = 1}^n a_2(X_i, Y_{2i}) \left(\frac{R_{2i}}{\pi_{2+}} -
      \frac{R_{1i}R_{2i}}{\pi_{11}}\right)\\
      &+ n^{-1} \sum_{i = 1}^n g_i \left(\frac{R_{1i}R_{2i}}{\pi_{11}} -
      1\right) \\
      &\equiv A + B + C + D.
    \end{align*}

  \item Notice that we will suppress the fact that response models are functions
    of $X$ (i.e. we write $\pi_{11}$ instead of $\pi_{11}(X)$).

  \item To compute the variance, we first solve for each covariance combination.
    Basically, all these computations rely of the following ideas. First, we
    assume that the response model is correctly specified. Consequently,
    $E[A] = E[B] = E[C] = E[D] = 0$ and things work out better. This helps when
    we take the covariance conditional on $X$ because the inner expectations are
    zero. The second key insight it to notice that $E[R_{j}^k] = E[R_j]$ for $j
    \in \{1, 2\}$ and $k \in \mathbb{N}$. This is because $R$ is a binary
    variable. Third, since we assume that the response models are correctly
    specified, we have $E[R_1 \mid X] = \pi_{1+}$, $E[R_2 \mid X] = \pi_{2+}$,
    and $E[R_1, R_2 \mid X] = \pi_{11}$.

    The overall approach to each of this computations is the following:
    (1) take conditional expectations with respect to $X$ (the $\Cov(E[\cdot])$
    term is zero), (2) expand the covariance to $E[XY] - E[X]E[Y]$ (the second
    term is also zero), (3) by the MAR assumption $g, a_2, b_2$ are independent
    of $R_1$ and $R_2$ and we can take the latter out of the expectation, (4)
    evaluate and simplify expressions involving $E[R]$.

    \begin{align*}
      \Cov(A, B) &= n^{-2} \sum_{i = 1}^n E\left[\Cov\left(E[g_i \mid X] \left(1
      - \frac{R_{1i}}{\pi_{1+}} - \frac{R_{2i}}{\pi_{2+}} +
      \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \mid X_i,\right. \right.\\
      &\qquad \qquad \qquad \qquad b_2(X_i, Y_{1i}) \left. \left.
      \left(\frac{R_{1i}}{\pi_{1+}} - \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \mid
      X_i\right)\right]\\
      &= n^{-1} E\left[E\left[E[g \mid X] \left(1 - \frac{R_{1i}}{\pi_{1+}} -
      \frac{R_{2i}}{\pi_{2+}} + \frac{R_{1i}R_{2i}}{\pi_{11}}\right)
      b_2(X_i, Y_{1i}) \left(\frac{R_{1i}}{\pi_{1+}} -
      \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \mid X \right]\right] \\
      &= n^{-1} E\left[E[g \mid X] E[b_2(X, Y_1) \mid X]
      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right].
    \end{align*}

    By symmetry,
    \begin{align*}
      \Cov(A, C) 
      &= n^{-1} E\left[E[g \mid X] E[a_2(X, Y_2) \mid X]
      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right].
    \end{align*}

    \begin{align*}
      \Cov(A, D) &= n^{-1} E\left[E\left[E[g \mid X] \left(1 -
      \frac{R_{1i}}{\pi_{1+}} - \frac{R_{2i}}{\pi_{2+}} +
      \frac{R_{1i}R_{2i}}{\pi_{11}}\right) g \left(\frac{R_1 R_2}{\pi_{11}} -
      1\right) \mid X\right] \right]\\
      &= n^{-1} E\left[E[g \mid X]^2 \left(\frac{-1}{\pi_{1+}} -
      \frac{1}{\pi_{2+}} + 2\right)\right].
    \end{align*}

    \begin{align*}
      \Cov(B, C) &= n^{-1} E\left[E[b_2(X, Y_1) \mid X] E[a_2(X, Y_2) \mid X]
      E\left[\left(\frac{R_1}{\pi_{1+}} - \frac{R_1 R_2}{\pi_{11}}\right)
      \left(\frac{R_2}{\pi_{2+}} - \frac{R_1 R_2}{\pi_{11}}\right) \mid
      X\right]\right]\\
      &= n^{-1} E\left[E[b_2(X, Y_1) \mid X] E[a_2(X, Y_2) \mid X]
      \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} - \frac{1}{\pi_{1+}} -
      \frac{1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right)\right].
    \end{align*}

    \begin{align*}
      \Cov(B, D) &= n^{-1} E\left[ E[b_2(X, Y_1) \mid X] E[g \mid X]
      E\left[\left(\frac{R_1}{\pi_{1+}} - \frac{R_1 R_2}{\pi_{11}}\right)
      \left(\frac{R_1 R_2}{\pi_{11}} - 1\right) \mid X\right] \right]\\
      &= n^{-1} E\left[ E[b_2(X, Y_1) \mid X] E[g \mid X]
      \left(\frac{1}{\pi_{1+}} - \frac{1}{\pi_{11}}\right)\right].
    \end{align*}

    By symmetry,
    \begin{align*}
      \Cov(C, D) 
      &= n^{-1} E\left[ E[a_2(X, Y_2) \mid X] E[g \mid X]
      \left(\frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}}\right)\right].
    \end{align*}

    We also compute the variance terms,

    \begin{align*}
      \Cov(A, A) &= n^{-1} E\left[ E\left[ E[g \mid X]^2 \left(1
      - \frac{R_{1i}}{\pi_{1+}} - \frac{R_{2i}}{\pi_{2+}} +
      \frac{R_{1i}R_{2i}}{\pi_{11}}\right)^2 \mid X\right] \right]\\
      &= n^{-1} E \left[ E[g \mid X]^2 \left(-1 + \frac{2 \pi_{11}}{\pi_{1+}
      \pi_{2+}} - \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
      \frac{1}{\pi_{11}}\right)\right].
    \end{align*}

    \begin{align*}
      \Cov(B, B) &= n^{-1} E\left[ E\left[b(X, Y_1)^2 \left(\frac{R_1}{\pi_{1+}}
      - \frac{R_1 R_2}{\pi_{11}}\right)^2 \mid X\right] \right]\\
      &= n^{-1} E\left[ E[b_2(X, Y_1)^2 \mid X] \left(\frac{-1}{\pi_{1+}} +
      \frac{1}{\pi_{11}}\right)\right].
    \end{align*}

    \begin{align*}
      \Cov(C, C) &= n^{-1} E\left[ E[a_2(X, Y_2)^2 \mid X]
      \left(\frac{-1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right) \right].
    \end{align*}

    \begin{align*}
      \Cov(D, D) &= n^{-1} E\left[ E\left[g_i^2 \left(\frac{R_1 R_2}{\pi_{11}} -
      1\right)^2 \mid X\right] \right]\\
      &= n^{-1} E\left[ E[g^2 \mid X] \left(\frac{1}{\pi_{11}} -
      1\right)\right].
    \end{align*}

    This means that
    \begin{align*}
      \Var(\hat \theta_{eff} - \hat \theta_n) 
      &= \Cov(A, A) + 2 \Cov(A, B) + 2 \Cov(A, C) + 2 \Cov(A, D) + \Cov(B, B) \\
      &+ 2 \Cov(B, C) + 2 \Cov(B, D) + \Cov(C, C) + 2 \Cov(C, D) + \Cov(D, D) \\
      &= n^{-1} E \left[ E[g \mid X]^2 \left(-1 + \frac{2 \pi_{11}}{\pi_{1+}
      \pi_{2+}} - \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
      \frac{1}{\pi_{11}}\right)\right] \\
      &+ 2 n^{-1} E\left[E[g \mid X] E[b_2(X, Y_1) \mid X]
      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right]\\
      &+ 2 n^{-1} E\left[E[g \mid X] E[a_2(X, Y_2) \mid X]
      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right]\\
      &+ 2 n^{-1} E\left[E[g \mid X]^2 \left(\frac{-1}{\pi_{1+}} -
      \frac{1}{\pi_{2+}} + 2\right)\right]\\
      &+ n^{-1} E\left[ E[b_2(X, Y_1)^2 \mid X] \left(\frac{-1}{\pi_{1+}} +
      \frac{1}{\pi_{11}}\right)\right]\\
      &+ 2 n^{-1} E\left[E[b_2(X, Y_1) \mid X] E[a_2(X, Y_2) \mid X]
      \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} - \frac{1}{\pi_{1+}} -
      \frac{1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right)\right]\\
      &+ 2 n^{-1} E\left[ E[b_2(X, Y_1) \mid X] E[g \mid X]
      \left(\frac{1}{\pi_{1+}} - \frac{1}{\pi_{11}}\right)\right]\\
      &+  n^{-1} E\left[ E[a_2(X, Y_2)^2 \mid X]
      \left(\frac{-1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right) \right]\\
      &+ 2 n^{-1} E\left[ E[a_2(X, Y_2) \mid X] E[g \mid X]
      \left(\frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}}\right)\right]\\
      &+ n^{-1} E\left[ E[g^2 \mid X] \left(\frac{1}{\pi_{11}} -
      1\right)\right].
    \end{align*}

    Differentiating yields:

    \begin{align*}
      \frac{\partial}{\partial a_2} \Var(\hat \theta_{eff} - \hat \theta_n)
      &= E\left[E[g \mid X] \left(\frac{1}{\pi_{1+}} + \frac{2}{\pi_{2+}} -
      \frac{2}{\pi_{11}} - \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right] \\
      &+ E\left[E[b_2(X, Y_1) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
      \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
      \frac{1}{\pi_{11}}\right)\right]\\
      &+ E\left[ E[a_2(X, Y_2) \mid X] \left(\frac{-1}{\pi_{2+}} +
      \frac{1}{\pi_{11}}\right)\right]\\
      &\equiv 0, \text{ and }\\
      \frac{\partial}{\partial b_2} \Var(\hat \theta_{eff} - \hat \theta_n)
      &= E\left[E[g \mid X] \left(\frac{2}{\pi_{1+}} + \frac{1}{\pi_{2+}} -
      \frac{2}{\pi_{11}} - \frac{\pi_{11}}{\pi_{1+}\pi_{2+}}\right)\right]\\
      &+ E\left[E[a_2(X, Y_2) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
      \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
      \frac{1}{\pi_{11}}\right)\right]\\
      &+ E\left[ E[b_2(X, Y_2) \mid X] \left(\frac{-1}{\pi_{1+}} +
      \frac{1}{\pi_{11}}\right)\right]\\
      &\equiv 0. 
    \end{align*}

    Substitution shows that these constraints are equivalent to:

    \begin{align*}
      E&\left[E[g \mid X] \left(\frac{-1}{\pi_{1+}} +
    \frac{1}{\pi_{2+}}\right)\right] + E\left[ E[b_2(X, Y_1) \mid X]
    \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -\frac{1}{\pi_{2+}}\right)\right]\\
      &- E\left[E[a_2(X, Y_2) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} - 
    \frac{1}{\pi_{1+}}\right)\right] \equiv 0
    \end{align*}

    where is the same as,

    \begin{align*}
      E&\left[ E[b_2(X, Y_1) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
    \frac{1}{\pi_{2+}}\right) \right] + E\left[E[g \mid X]
      \left(\frac{1}{\pi_{2+}}\right)\right] \\
      &=
    E\left[ E[a_2(X, Y_2) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
    \frac{1}{\pi_{1+}}\right) \right] + E\left[E[g \mid X]
    \left(\frac{1}{\pi_{1+}}\right)\right].
    \end{align*}

  \item These constraints can be satisfied (this is sufficient but maybe not
    necessary) if

    \begin{align*}
      E&\left[(E[b_2(X, Y_1) \mid X] - E[a_2(X, Y_2) \mid X])
      \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right) \right] = 0\\
      E&\left[\left(\frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}}\right)(E[a_2(X, Y_2)
      \mid X] + E[b_2(X, Y_1) \mid X] - 2E[g \mid X])\right] = 0.
    \end{align*}

@@

\textbf{Questions for Dr. Kim}
\begin{itemize}
  \item Where do we go from here?
  \item This is a functional expectation that I need to calibrate. Should I use
    some sort of basis spline to try to estimate these expectations and then set
    them equal to zero? I don't think that I have worked with this kind of
    constraint before.
\end{itemize}
%\input{prop_full_nonmonoest.tex}
